---
title: "Human Activity Recognition Prediction"
author: "Vincent Dupont"
date: "8/15/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(caret)
library(gbm)
library(rpart)
library(rpart.plot)
library(rattle)
set.seed(1234)
```

## Executive Summary

The purpose of this prediction study design is to be able to answer: In which position (class feature in the dataset) was the subject when doing the activity?


## Data

To answer the question we are going to use the Human Activity Recognition dataset that can be found at the following url:

http://groupware.les.inf.puc-rio.br/har

### Loading

```{r }
data <- read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", header=T)
validationData <- read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", header=T)
dim(data)
```

### Cleaning

#### Removing near zero variance variables
```{r }
NZV <- nearZeroVar(data)
data <- data[, -NZV]
dim(data)
```

#### Removing NA
```{r }
anyNA    <- sapply(data, function(x) any(is.na(x)))
data <- data[, anyNA==FALSE]
dim(data)
```

#### Removing useless features
```{r }
data <- within(data, rm(X, user_name, raw_timestamp_part_1, raw_timestamp_part_2, cvtd_timestamp, num_window))
dim(data)
```

### Splitting
```{r }
inTrain <- createDataPartition(data$classe, p = 0.7, list = FALSE)
trainingData <- data[inTrain, ]
testData <- data[-inTrain, ]
```

## Algorithms

### Random Forest
```{r, cache=TRUE}
trControlRF <- trainControl(method="cv", number=2, verboseIter=FALSE)
fitRF <- train(classe ~ ., data=trainingData, method="rf",
                          trControl=trControlRF)
fitRF$finalModel

```

```{r}
predictRF <- predict(fitRF, newdata=testData)
confusionMatrix(predictRF, testData$classe)
```

Random Forest Accuracy : 0.9941

### Decision Tree
```{r, cache=TRUE}
trControlDT <- trainControl(method="cv", number=5)
fitDT <- train(classe~., data=trainingData, method="rpart", trControl=trControlDT)
fitDT$finalModel
```

```{r}
fancyRpartPlot(fitDT$finalModel)
```


```{r}
predictDT <- predict(fitDT, newdata=testData)
confusionMatrix(predictDT, testData$classe)
```

Decision Tree Accuracy : 0.489

### Generalized Boosted Model

```{r, cache=TRUE }
trControlGbm <- trainControl(## 10-fold CV
                           method = "repeatedcv",
                           number = 5,
                           ## repeated two times
                           repeats = 2)

fitGbm <- train(classe ~ ., data = trainingData,
                  method = "gbm",
                  trControl = trControlGbm,
                  verbose = FALSE)
fitGbm$finalModel
```

```{r}
predictGBM <- predict(fitGbm, newdata=testData)
confusionMatrix(predictGBM, testData$classe)
```

Generalized Boosted Model Accuracy : 0.9645


## Evaluation

The algorithm with the highest accuracy is Random Forest with 0.9941.

In our final step we are going to use this algorithm with our valication set.

```{r}
predictions <- predict(fitRF, newdata=validationData)
data.frame(predictions=predictions)
```
